---
title: "Sales Data Analysis"
format: beamer
editor: visual
echo: false
---

```{r}
library(readxl)
library(tidyverse)
library(ggplot2)

SalesData <- read_excel("SalesData.xlsx")
```

## Introduction

-   This presentation analyzes the *SalesData* dataset.
-   The dataset contains sales information across:
    -   Multiple **years**
    -   Different **cities**
    -   Various **product categories**
-   The objective is to identify:
    -   Year-wise sales trends
    -   City-wise performance patterns

## Year-wise Sales

```{r, fig.height=2,fig.width=6}
year_sales <- tapply(SalesData$Sales, SalesData$Year, sum)
yearwise_sales <- data.frame(
  Year = names(year_sales),
  Total_Sales = as.numeric(year_sales)
)
knitr::kable(yearwise_sales)
yearwise_sales$Year <- as.numeric(yearwise_sales$Year)

ggplot(yearwise_sales, aes(Year, Total_Sales)) +
  geom_line()
```

Sales show a steady year-on-year increase from 2021 to 2025, indicating consistent business growth and stable performance.

## City-wise Sales Comparison

```{r,fig.height=4}
city_sales <- tapply(SalesData$Sales, SalesData$City, sum)

city_sales_df <- data.frame(
  City = names(city_sales),
  Total_Sales = as.numeric(city_sales)
)

ggplot(city_sales_df, aes(City, Total_Sales, fill = City)) +
  geom_col()
```

Sales vary significantly across cities, with Delhi and Mumbai leading in total sales, while Ahmedabad and Kolkata show lower contributions.

## Operating Profit Across Product Categories

```{r,fig.height=4}
ggplot(SalesData, aes(Category, Operating_Profit, fill = Category)) +
  geom_boxplot()
```

Product-wise profit analysis shows that Personal Care and Home Care are the most profitable categories, indicating stronger margins and operational efficiency.
